Approximation as Differentiation

### 2025-05-05 appendix/notes/auto-diff.tex
Category Diff.

For any map $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$, write $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$. We can define composition on such maps as:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent and verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
\linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and
$(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.

>>>>

Write $\Diff$ for the category where the objects are all finite vector spaces $\RR^n$ and the morphisms are
all maps $f: \RR^m \to \RR^n \times (\RR^m \linearto \RR^n)$. Writing $f_1$ for $\pi_1 \comp f$ and $f_2$ for
$\pi_2 \comp f$, composition in $\Diff$ is given by:
\begin{align*}
(g \comp f)_1(x) &= f_1(g_1(x)) \\
(g \comp f)_2(x) &= g_2(f_1(x)) \comp f_2(x)
\end{align*}
\noindent We can verify that $\tangents$ is indeed a functor, regardless of whether the linear maps $\RR^m
\linearto \RR^n$ happen to be derivatives. But if $f_2(x)$ happens to be the derivative of $f_1$ at $x$ for
any $x \in \RR^m$ and similarly for $g$ then the derivatives compose according to the forward chain rule and
$(g \comp f)_2(x)$ is the derivative of $(g \comp f)_1$ at $x$.
